Convergence rate of MAP estimates \\ for the exponential family

### 2021-09-30 main.tex
Finish proof with exponent alpha

\subsection{Other Idea}
There are two main bottlenecks when casting MAP as an iterative algorithm. 
\begin{enumerate}
	\item The first one is the variance definition which always end up being infinite when asked to be uniform on every possible trajectories, if the log-partition  has some barrier structure.
	\item The second one is the step-size scheduling. $O(\inv{t})$ is decreasing too fast for most proof techniques. Even in the Euclidean setting where the variance does not matter, it results in a disappointing rate $O(\inv{\log t})$.
\end{enumerate}
Regarding this second point, it might be possible to do something comparable to \citet{gower2019sgd} who get a $O(\inv{t})$ rate with a $O(\inv{t})$ step-size. Take $m$ and $L$ as the (relative) strong-convexity and smoothness constants of the problem.Their descent lemma looks like
\begin{align}
	B_{t+1} \leq ( 1 - \lr_t m) B_t + \lr_t^2 \sigma^2
\end{align}
where $B_t$ is some measure of suboptimality, in this case the $\ell^2$ distance between $\natp^t$ and $\natp^*$ ; and $\lr_t \leq \inv{2L}$.
On the other hand, the problem of most Bregman analysis is that they do not get this $\lr^2$ in front the variance. Instead they simply get a $\lr$, because they do not benefit from the homogeneity of the norm
\begin{align}
	B_{t+1} \leq ( 1 - \lr_t m) B_t + \lr_t \sigma^2 \; .
\end{align}
The problem is that this exponent $2$ is crucial to get fast convergence with a $O(\inv{t})$ step-size. Otherwise the variance is too large and we need to decrease the step-size more slowly to reach the optimum quickly.
Now our question is what happens if we get something in-between ? e.g.
\begin{align}
	B_{t+1} \leq ( 1 - \lr_t m) B_t + \lr_t^\alpha \sigma^2 \; .
\end{align}
where $1<\alpha<2$ is some exponent characterizing the geometry of the problem.
Inspired by \citet{gower2019sgd} we set
\begin{align}
	\lr_t := \inv{m} \left ( 1 - \left (\frac{k}{k+1} \right)^\alpha \right ) \; .
\end{align}

>>>>

\subsection{Other Idea}
There are two main bottlenecks when casting MAP as an iterative algorithm. 
\begin{enumerate}
	\item The first one is the variance definition which always end up being infinite when asked to be uniform on every possible trajectories, if the log-partition  has some barrier structure.
	\item The second one is the step-size scheduling. $O(\inv{t})$ is decreasing too fast for most proof techniques. Even in the Euclidean setting where the variance does not matter, it results in a disappointing rate $O(\inv{\log t})$.
\end{enumerate}
Regarding this second point, it might be possible to do something comparable to \citet{gower2019sgd} who get a $O(\inv{t})$ rate with a $O(\inv{t})$ step-size. Take $m$ and $L$ as the (relative) strong-convexity and smoothness constants of the problem, along with $K=\frac{L}{m}$ the condition number.Their descent lemma looks like
\begin{align}
	B_{t+1} \leq ( 1 - \lr_t m) B_t + \lr_t^2 \sigma^2
\end{align}
where $B_t$ is some measure of suboptimality, in this case the $\ell^2$ distance between $\natp^t$ and $\natp^*$ ; and $\lr_t \leq \inv{2L}$.
On the other hand, the problem of most Bregman analysis is that they do not get this $\lr^2$ in front the variance. Instead they simply get a $\lr$, because they do not benefit from the homogeneity of the norm
\begin{align}
	B_{t+1} \leq ( 1 - \lr_t m) B_t + \lr_t \sigma^2 \; .
\end{align}
The problem is that this exponent $2$ is crucial to get fast convergence with a $O(\inv{t})$ step-size. Otherwise the variance is too large and we need to decrease the step-size more slowly to reach the optimum quickly.
Now our question is what happens if we get something in-between ? e.g.
\begin{align}
	B_{t+1} \leq ( 1 - \lr_t m) B_t + \lr_t^\alpha \sigma^2 \; .
\end{align}
where $1<\alpha<2$ is some exponent characterizing the geometry of the problem.

\paragraph{Step-size.}
Inspired by \citet{gower2019sgd} we set
\begin{align}
	\lr_t := \inv{m} \left ( 1 - \left (\frac{t}{t+1} \right)^\alpha \right ) \; .
\end{align}
The function $x\mapsto (1+x)^\beta$ is convex, meaning that 
\begin{align}
	\left (\frac{t}{t+1} \right)^\alpha 
	= \left (1 - \frac{1}{t+1} \right)^\alpha 
	\gtrsim 1 - \frac{\alpha}{t+1}
\end{align}
where the $\gtrsim$ symbol stands for greater or equal $\geq$ and asymptotically equivalent $\sim$.
Back to the step-size, we get
\begin{align}
	\lr_t \lesssim \frac{\alpha}{m(t+1)} \; .
	\label{eq:lr_asymptote}
\end{align} 
From this we draw two conclusions.
First the step-size is in the right asymptotic category $\lr_t \in \Theta(t^{-1})$.
Second, we have a sufficient condition for the step-size to be small enough $\lr_t \leq \inv{2L}$ and the descent lemma to apply
\begin{align}
	\lr_t \leq \inv{2L}
	\impliedby \frac{\alpha}{m(t+1)}  \leq \inv{2L} 
	\iff \frac{2 \alpha L}{m} \leq t +1
	\impliedby t \geq t_0 := \ceil{2\alpha K} -1
\end{align}
where $K= L/m$ is the condition number of the problem.
Consequently, when $t < t_0 :== \ceil{2\alpha K}$, we set $\lr_t = \inv{2L}$.
Let us correct our definition to account for this revision
\begin{equation}
	\lr_t := \begin{cases}
		\inv{2L} \ \text{ if } \ t< t_0 = \ceil{2\alpha K} -1, \\
		\inv{m} \left ( 1 - \left (\frac{t}{t+1} \right)^\alpha \right ) \ \text{otherwise.}
	\end{cases}
\end{equation}

\paragraph{Telescopic Sum.}
For $t \geq t_0$, the step-size definition means that the decreasing factor of the descent lemma is
\begin{align}
	1- \lr_t m = \left(\frac{t}{t+1} \right)^\alpha  \; .
\end{align}
Multiplying both sides of the descent lemma with $(t+1)^\alpha$ gives
\begin{align}
	(t+1)^\alpha B_{t+1} 
	\leq t^\alpha B_t + ((t+1)\lr_t)^\alpha \sigma^2 \; .
\end{align}
Thanks to~\eqref{eq:lr_asymptote}, we know that $(t+1)\lr_t \leq \frac{\alpha}{m}$, thus the second term or the upper bound is upper bounded as
\begin{align}
	((t+1)\lr_t)^\alpha \sigma^2  
	\leq \left(\frac{\alpha}{m} \right)^\alpha  \sigma^2 \; .
\end{align}
By recursion (or telescopic summing), we get
\begin{align}
	t^\alpha B_t \leq t_0^\alpha B_{t_0} 
	+ (t - t_0) \left(\frac{\alpha}{m} \right)^\alpha  \sigma^2  \; .
\end{align}
It is possible to get the constant step-size rate for this $B_{t_0}$, but for now, we already see that
\begin{align}
	B_t \leq \left(\frac{t_0}{t} \right)^\alpha B_{t_0} 
	+ \frac{t - t_0}{t^\alpha} \left(\frac{\alpha}{m} \right)^\alpha  \sigma^2 \in \Theta(t^{1-\alpha}) 
\end{align}
where $-1 <1-\alpha < 0$ and this slowly decreasing factor is coming from the variance term.
